---
title: "Income and Right to Work in the United States"
output: html_document
group: Evan L, Evan M, Darcy, Ana, and Caz
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warnings = FALSE)
#Pulling packages needed for initial analysis
install.packages("pacman", repos = "http://cran.us.r-project.org")
library("pacman")
pacman::p_load("haven", "ggplot2", "scales", 
               "dplyr", "magrittr", "ggthemes")
```

# Dataset
###### We are drawing our data from a subsample of the American Community Survey (ACS) data from the year 2016. The ACS is a survey conducted annually by the U.S. Census Bureau, gathering long-form information previously collected in the Decennial (conducted in years ending 0) Census. 

# Policy Analysis
######  Right to Work (RTW) refers to state laws that prohibit unions from compelling individuals to join, or pay union dues as employment condition. In addition, right to work laws sometimes also prohibit contracts that stipulate only hiring unionized workers. 

###### Although Right to Work laws may seem beneficial,  critics argue that they are inherently "anti-union". There are already federal laws existing that prevent mandatory union membership, and some argue that RTW laws harm union's bargaining strength, effectively lowering salaries and benefits. 

## Purpose of Analysis
###### Our analysis will seek to address the aforementioned issues, and analyze overall income differences in the US.
  + Do states with RTW laws have lower income?
  + Is there a significant difference in income of men vs women, via RTW   laws?


```{r}
dataset <- 'https://github.com/EvanLih/PUBPOL599_Right_To_Work/raw/master/Dataset/Cleaned_RTW_Dataset.csv'
dataset <- read.csv(dataset, stringsAsFactors = F) 
dataset$Right_to_Work %<>% as.factor
```

```{r}
#converting the Right_to_Work column into as a factor, as it is currently being read as a integer. 
means <- aggregate(incwage ~ Right_to_Work, dataset, mean)
means$incwage %<>% round(2)

#I use the "suppressMessages" command so that the errors generated by ggplot are not printed out in our markdown file. However, because ggplot does not generate messages, we need to make sure to include the "print" command. 
suppressMessages(print(ggplot(dataset, aes(Right_to_Work, incwage, fill = Right_to_Work, show.legend = FALSE), group = Right_to_Work) +
  geom_boxplot(outlier.color = "red", outlier.shape = NA) +
  scale_y_continuous(breaks = c(10000,30000,50000,70000,100000), labels = dollar) +
  ylim(0,150000) +
  labs(x = "Right to Work", y = "Income (Dollars)") +
  geom_text(data=means, aes(label = incwage, y = incwage + .09)) +
  scale_x_discrete(breaks= c(0,1), labels = c("No Right to Work", "Right to Work")) +
  scale_fill_manual(values = c("dodgerblue4", "indianred4")) +
  labs(title = "Median Income by Right to Work Status") +
  theme_stata() +
  scale_color_stata()+
  theme(plot.subtitle=element_text(size=10, hjust=0.5, color="black")) +
  theme(plot.title=element_text(size=19, face="bold", hjust = .5)) +
  theme(axis.title.x=element_text(face = "bold", size = 10)) +
  theme(axis.title.y = element_text(face = "bold", size = 10)) +
  theme(axis.text.y = element_text(angle = 45)) +
  theme(legend.title=element_blank()) + 
  guides(fill=FALSE)))

LinRegwage = lm(incwage ~ Right_to_Work, data = dataset)
summary(LinRegwage)

```
#Our results are **Significant**;
#####On average, all else equal, being an RTW State is associated with a decrease of USD 7736.112 in the wage income. Also, on average, all else equal, one more year in age is associated with an increase of approximately USD 71 in annual wages compared to being a woman. 



```{r}
ggplot(dataset, aes(Right_to_Work, incwage, fill = Right_to_Work), group = Right_to_Work) +
  geom_bar(stat = "summary", fun.y = "mean", position = position_dodge(width = 2.0), width = .7) +
  labs(x = "Right to Work", y = "Income (dollars)") +
  scale_x_discrete(breaks= c(0,1), labels = c("No RTW", "Has RTW")) +
  scale_y_continuous(breaks = c(10000,20000,30000,40000,50000,60000), labels = dollar) +
  scale_fill_manual(name= "Right to Work Status", labels=c("No Right to Work", "Right to Work"),
                    values = c("dodgerblue4", "indianred4")) +
  facet_grid(. ~ sex) +
  labs(title = "Average Income by Sex", subtitle = "Divided by Right to Work/ No Right to Work") +
  theme_stata() +
  scale_color_stata()+
  theme(plot.subtitle=element_text(size=10, hjust=0.5, color="black")) +
  theme(plot.title=element_text(size=19, face="bold", hjust = .5)) +
  theme(axis.title.x=element_text(face = "bold", size = 10)) +
  theme(axis.title.y = element_text(face = "bold", size = 10)) +
  theme(axis.text.y = element_text(angle = 45))

incSexMeans <- dataset %>% group_by(sex, Right_to_Work) %>% summarise(mean = mean(incwage))

print(incSexMeans)

```

For Females in states with no RTW laws, the average income is __35929__, while females in states with RTW laws earn __29113__. 
For Males in states with no RTW laws, the average income is __57619__, while males in states with RTW earn __49021__. 
These results shows that the presence of RTW laws effect both male and female income earnings
```{r}
incSex <- t.test(incwage~sex, data = dataset)
print(incSex)
```
This t-test shows if there is a significant difference in the mean values of income between Males and Females. With a P-value of 2.2e-16, we can safely reject the null hypothesis that there is no difference in the mean values of income between Males and Females. 

```{r}
testRTW=lm(as.numeric(incwage)~Right_to_Work+age,data=dataset)
summary(testRTW)
```

Our adjusted R-squared for this regression is small...
```{r}
summary(testRTW)$adj.r.squared
```

```{r}
ggplot(dataset, aes(age, incwage, show.legend=TRUE)) +
  stat_summary(fun.y="mean",geom="point")+
  labs(x = "Age", y = "Income (dollars)") +
  scale_y_continuous(breaks=seq(10000,60000,by=10000), labels = dollar)+
  facet_grid(. ~ Right_to_Work)+
  labs(title="Average Income by Age by RTW Status", caption = "IPUMS", subtitle = "0=No RTW, 1= RTW") +
  theme_stata() +
  scale_color_stata()+
  theme(plot.subtitle=element_text(size=10, hjust=0.5, color="black")) +
  theme(plot.title=element_text(size=13, face="bold", hjust = .5)) +
  theme(axis.title.x=element_text(face = "bold", size = 10)) +
  theme(axis.title.y = element_text(face = "bold", size = 10)) +
  theme(axis.text.y = element_text(angle = 45))
```
```{r}
t.test(age~Right_to_Work, data = dataset)
test = glm(Right_to_Work ~ age, data = dataset, family = binomial())
t.test(incwage~Right_to_Work, data = dataset)

test = glm(Right_to_Work ~ incwage, data = dataset, family = binomial())
```

